Query Optimization for Dynamic Imputation

### 2017-05-15 paper/rebuttal.tex

  \item S4.4.1: Imputation over the raw data is being used as ground truth. Is there any reason to suspect that this is the correct thing to do? Perhaps joining multiple tables together exposes new attributes that makes imputation *more* reliable? At the very least, it would be useful to understand why there is so much error in $\alpha=0$ on Q2 and Q8.

    \resp{For our purposes, imputation over the raw data is a reasonable ground truth. This
      is because the analyst, in practice, should be considering the tradeoffs on quality
      and performance of using ImputeDB versus a production imputation system on the raw
      data. Unfortunately, they would never have access to real ground truth in such a
      scenario, and moreover, finding ground truth for use even in an empirical evaluation
      of imputation methods can be challenging because artificially dirtying the data
      often does not capture the complex reasons for missingness in real world datasets.
      We updated the discussion in \Cref{sec:results} to emphasize this point.}

  \item There are a number of opportunities for trimming in the figures, tables, and algorithms: Algorithms 1-3 could be compacted a bunch, Table 1 is full of whitespace (e.g., make it a 3-column {figure*}), Tables 1-2 could also be summarized as CDFs, Figure 6 is full of whitespace, Table 3 could be put queries on a single row (e.g., Q1 would easily fit on 1 line), and the references have a lot of redundant information (URLs, for example).

    \resp{We have compacted the algorithms, tables, and references. Note that Tables 1 and 2 show the percentage of missing values for \emph{each attribute}, not as an overall percentage, so cannot be summarized as CDFs. We think that the URLs add value to the references, particularly for data sets.}
  \end{itemize}

@@

  \item D5. Can you have plans with a mix of drop and impute operators?

    \resp{Yes, we have now highlighted this for clarity in the paper as well
    (\Cref{fig:example-plans} and \Cref{sec:search-space})}

  \item D6. During plan enumeration, can you safely prune plans at some point, similar to what a traditional optimizer would do?

    \resp{We maintain the set of Pareto-optimal plans at each point in the optimization. This set is significantly smaller than the set of all plans, so it represents some pruning (a point we have
    now clarified in the paper in \Cref{subsubsec:pareto}). We found that the size of these Pareto sets is small enough that we did not need to prune further.

      If necessary, it is possible to sample from the Pareto-optimal plans during the plan
      enumeration. The final selected plan may not be Pareto-optimal, but this is a potential way to control the number of plans.
      To perform this approximation, we only add plans to the frontier that are non-dominated and are sufficiently different from existing plans in the frontier. The plan already in the frontier should
      provide similar tradeoffs and reduces the search space explored. \ProjectName{} only employs this optimization when a large number of tables requiring imputation are involved in the query.
      }

### 2017-05-15 paper/conclusion.tex
paper: temp widow fix
Incorporate Nicola's comments on draft

\section{Conclusion}
Our evaluation of \ProjectName{} shows that imputation of missing values can be successfully integrated into the relational algebra and
existing cost-based plan optimization frameworks. We implement imputation actions, such as dropping or imputing values with a machine
learning technique, as operators in the algebra and use a simple yet effective cost model to consider trade-offs in 
imputation quality and runtime. We show that different values for the trade-off parameter can yield substantially
different plans, allowing the user to express their preferences for performance on a per-query basis.
Our experiments with the CDC NHANES, freeCodeCamp and ACS datasets
show that \ProjectName{} can be used successfully with real-world data to run standard SQL queries over data with missing values. 
We craft a series of realistic queries, meant to resemble the kind of questions an analyst might ask during
the data exploration phase. The plans selected for each query execute an order-of-magnitude faster than
 the standard approach of imputing on the entire base tables and
then formulating queries. Furthermore, the difference in query results between the two approaches is
shown to be small in all queries considered (\lowsmapealphazero{}-\highsmapealphaone{}\% error, \Cref{sec:results}). Data analysts need not commit to a specific imputation strategy and can instead
vary this across queries.

@@

\subsection{Future Work}
\ProjectName{} opens up multiple avenues for further work.

>>>>

\section{Conclusion}
Our implementation of \ProjectName{} and experiments show that imputation of missing values can be successfully integrated into the relational algebra and
existing cost-based plan optimization frameworks. We implement imputation actions, such as dropping or imputing values with a machine
learning technique, as operators in the algebra and use a simple, yet effective, cost model to consider trade-offs in 
imputation quality and runtime. We show that different values for the trade-off parameter can yield substantially
different plans, allowing the user to express their preferences for performance on a per-query basis.
Our experiments with the CDC NHANES, freeCodeCamp and ACS datasets
show that \ProjectName{} can be used successfully with real-world data, running standard SQL
queries over tuples with missing values. 
We craft a series of realistic queries, meant to resemble the kind of questions an analyst might ask during
the data exploration phase. The plans selected for each query execute an order-of-magnitude faster than
 the standard approach of imputing on the entire base tables and
then formulating queries. Furthermore, the difference in query results between the two approaches is
shown to be small in all queries considered (\lowsmapealphazero{}-\highsmapealphaone{}\% error, \Cref{sec:results}). Data analysts need not commit to a specific imputation strategy and can instead
vary this across queries.

@@

\pagebreak
\subsection{Future Work}
\ProjectName{} opens up multiple avenues for further work.
